### 2015-02-14 doc/meeting.Rmd

---
output:
  pdf_document: default
---
Meeting at INESC, on 2014-06-14
=====
Price data
----
- pick some stocks and check 
- get price form data stream ( iterate )

Black-Litterman
--
## Strategy
- create true from non ranking (at t+1)
- same omega, different views

## Story: paper structure
build on top of ssrn paper

### 1. use paper and reproduce strategy

- {views: meanTPER from t-1; conf: coef of variation}

### 2. _Rank matters_
- {views: baseline rankings t-1 (ranking on PT); conf=coef. of variation}

- {views: true ranking(PT), meanTPER from t; conf: 1}

### 3. Last part
- {views: true ranking (EPS), meanTPER from t; conf: 1, coef of variation}

## result assumptions:
- ranking god
- average god
- naive
- default

## Constrains
Run the code with no short-sale

-----

Meeting with Ana Paula 2014-07-28
====
## More research DS PT vs. our PT
1. Check the min and max of $PT/DS.PT$ ratio
2. Contact Thompson/DataStream for help
3. Check AMZN price targets over the net and compare.