### 2015-02-14 doc/paper/paper-test.Rmd

---
title: "Paper outline"
output:
  pdf_document: default
  word_document: default
---
## Possible title
"Are rankings of financial analysts useful for investors?"

### 1. Introduction:
  + Rankings are important;
  + Who also ranks analysts;
  + Paper objectives:
    + operationalize BL;
    + ranking based strategy outperforms baselines;
  + Paper structure;

### 2. Rankings of FA:
  + Efficient Market Hypothesis and advice of FA;
  + Can we make money if EMH holds?
  + FA do have skills;
  + Rankings of FA by StarMine:
    + based on recommendations (target prices);
    + based on EPS forecast accuracy (SES metrics);
  + FA rankings signal who to follow;
  + Our FA rankings methods:
    + naive;
    + default;
    + non ranking benchmark;

### 3. Trading strategy:
  + Why BL and not CAMP?;
  + BL notations;
  + BL strategy outline;
  + Defining expected returns:
    + rank-based;
    + non-rank based (mean);
  + Defining the confidence (coefficient of variation);

### 4. Data and experiment setup:
  + Data description (sources, total observations);
  + TPER is truncated;
  + Descriptive stats of filtered data;
  + How we rank analysts?
  + How we create:
    + naive ranking;
    + default ranking;

### 5. Results
  + BL strategies outperform market;
  + _true_ of rank based outperforms non-rank;
  + Foreknowledge of rankings leads to highest cumulative return;
  + FA Rankings are important, goal is to forecast them correctly;

### 6. Conclusion
  + Ranking of FA are important and produce the highest possible cumulative return;
  + BL strategy takes rankings as inputs to build views;
  + Future research: predict rankings of FA;

### References

### Tables and Plots