### 2015-02-14 doc/trading.Rmd

```{r echo=FALSE, warning=FALSE}
rm(list=ls())
setwd('~/Documents/Dropbox/workspace/Projects/Black-Litterman/BL-strategies/')
library(ProjectTemplate)
library(scales)
load.project()
delta<-2.5
```

``` {r echo=FALSE, cache=TRUE, warning=FALSE}

#source('munge/01-data.pre.process.R')
###subset of core stocks/broker: min. 3 broker in quarter and min. 12 quarters of coverage 
q.data=q.data[,core.b:=.N>11,by=list(Stock,Broker)]
core.dt <- q.data[(core.b)][,rank:=rank(score),by=list(q.id,Stock)]
core.dt=core.dt[,clean.s:=.N>2,by=list(q.id,Stock)]
###core rankings
pt.new <- acast(core.dt[(clean.s)],q.id~Broker~Stock,value.var='rank')
###Ranking: making brokers view (extended TPER for 4 quarters)
brok.tper <- acast(unique(ext.pt,by=c('Stock','V2','Broker')),V2~Broker~Stock,value.var='b.view')

###an alternaive to measure accuracy: coef. of variation. To be used when compareing with non ranking methods
brok.cv <- acast(unique(q.data,by=c('Stock','q.id')),q.id~Stock~Broker,value.var='b.coefVar')

##BL inputs for non-rank strateby:meanTper and s.coefVar
meanTper <-acast(unique(q.data,by=c('Stock','q.id')),q.id~Stock,value.var='meanTper')

###Coefficient of variation: lower values means less variable,thus, more confidence. The higher the CV, the greater the dispersion in the variable [http://www.ats.ucla.edu/stat/mult_pkg/faq/general/coefficient_of_variation.htm]
conf.coef <- acast(unique(q.data,by=c('Stock','q.id')),q.id~Stock,value.var='s.coefVar')
### rescale the CV to (0,1). To be consistent with the ranking model, also scale the CV so that if CV=0, don't trust and if CV=1 trust
require(scales)
norm.conf<- rescale(1-conf.coef[1:44,])

#methods<-c('raw','1diff','random','roll.sd')
#baselines<-c('true','naive','default',methods)
#ranking <- 'pt.new'
percentile <- c(0.01,0.99)
trunk.meanTper <- truncate.f(meanTper[1:44,],percentile)
all.s <- findIntersect(pt.new,brok.tper,3)
all.b <- findIntersect(pt.new,brok.tper,2)

trunk.brok.tper <- truncate.f(brok.tper[1:dim(pt.new)[1],all.b,all.s],percentile)

true.Q <- buildQs(all.s,pt.new,trunk.brok.tper)
true.conf <- true.Q
true.conf[!is.na(true.conf)] <- 1

plot.true <- ggplot(data.table(melt(true.Q,na.rm=T)),aes(x=as.Date(as.yearqtr(Var1)),y=value,group=Var1))+geom_boxplot()+theme_bw()
plot.conf <- ggplot(data.table(melt(norm.conf,na.rm=T)),aes(x=as.Date(as.yearqtr(Var1)),y=value,group=Var1))+geom_boxplot()
```
``` {r}
plot.true
plot.conf
```

```{r samePeriod, echo=FALSE}
bl.period <- 3:(dim(trunk.meanTper)[1])
m.period <- ((length(market.list)-length(bl.period))+1) : (length(market.list))

#opt.w<- rbindlist(script.weights.env.f('meanTper',market.list[m.period],trunk.meanTper[bl.period,],norm.conf[bl.period,],0.02,2.5))

opt.w<- rbindlist(script.weights.env.f('rankTper',market.list[m.period],true.Q[bl.period,],true.conf[bl.period,],0.02,2.5))

bl.results <- unique(opt.w,by='q.id')
bl.results=bl.results[,c('cum.ret','ann.ret','ann.sd','meanViews'):=cum.ret.f(port.ret,n.views)][,ann.sr:=ann.ret/ann.sd]
```
# view and market in the same period: start p = 3, ann.ret=`r bl.results[1,ann.ret]`, sr=`r bl.results[1,ann.sr]`
```{r p.samePeriod,echo=FALSE}
plot(bl.results[,cum.ret],type='l')
```

```{r viewPlus, echo=FALSE}
bl.period <- 3:(dim(trunk.meanTper)[1])
m.period <- ((length(market.list)-length(bl.period))) : (length(market.list)-1)

#opt.w<- rbindlist(script.weights.env.f('meanTper',market.list[m.period],trunk.meanTper[bl.period,],norm.conf[bl.period,],0.02,2.5))

opt.w<- rbindlist(script.weights.env.f('rankTper',market.list[m.period],true.Q[bl.period,],true.conf[bl.period,],0.02,2.5))

bl.results <- unique(opt.w,by='q.id')
bl.results=bl.results[,c('cum.ret','ann.ret','ann.sd','meanViews'):=cum.ret.f(port.ret,n.views)][,ann.sr:=ann.ret/ann.sd]
```
# view one period ahead of market: start p = 3, ann.ret=`r bl.results[1,ann.ret]`, sr=`r bl.results[1,ann.sr]`
```{r p.viewPlus,echo=FALSE}
plot(bl.results[,cum.ret],type='l')
```

```{r viewMinus, echo=FALSE}
bl.period <- 3:(dim(trunk.meanTper)[1]-1)
m.period <- ((length(market.list)-length(bl.period))+1) : (length(market.list))

#opt.w<- rbindlist(script.weights.env.f('meanTper',market.list[m.period],trunk.meanTper[bl.period,],norm.conf[bl.period,],0.02,2.5))

opt.w<- rbindlist(script.weights.env.f('rankTper',market.list[m.period],true.Q[bl.period,],true.conf[bl.period,],0.02,2.5))

bl.results <- unique(opt.w,by='q.id')
bl.results=bl.results[,c('cum.ret','ann.ret','ann.sd','meanViews'):=cum.ret.f(port.ret,n.views)][,ann.sr:=ann.ret/ann.sd]
```
# view behind maket one period: ann.ret=`r bl.results[1,ann.ret]`, sr=`r bl.results[1,ann.sr]`
```{r p.viewMinus,echo=FALSE}
plot(bl.results[,cum.ret],type='l')
```

```{r ToDo,echo=FALSE}
# bl.results[,TO:=data.table(c(NA,sapply(2:42,function(b){
# beg <- setkey(opt.w[q.id==b,list(Stock,opt.w)],Stock)
# end <- setkey(opt.w[q.id==b-1,list(Stock,opt.w)],Stock)
# end[beg][,turnover.f(opt.w,opt.w.1)]})))]
# 
# 
# 
# 
# 
# all.s <- findIntersect(get(paste(ranking,'all.rankings',sep='.')),data,3)
# all.b <- findIntersect(get(paste(ranking,'all.rankings',sep='.')),data,2)
# #all.s <- intersect(intersect(intersect(dimnames(pt.all.rankings)[[3]],dimnames(eps.all.rankings)[[3]]),dimnames(data)[[3]]),dimnames(eps.all.rankings)[[3]])
# 
# #all.b <- intersect(intersect(intersect(dimnames(pt.all.rankings)[[2]],dimnames(eps.all.rankings)[[2]]),dimnames(data)[[2]]),dimnames(eps.all.rankings)[[2]])
# 
# #all.rankings <- abind(pt=pt.all.rankings[2:42,all.b,all.s,,1],eps=eps.all.rankings[,all.b,all.s,,1],along=5)
# 
# all.rankings <- abind(pt.new=get(paste(ranking,'all.rankings',sep='.'))[,all.b,all.s,,1],along=5)
# 
# accuracy <- abind(pt.new=get(paste(ranking,'accuracy',sep='.'))[,,all.s,1],along=4)
# 
# #accuracy <- abind(eps=eps.accuracy[,2:42,all.s,1],pt=pt.accuracy[,2:42,all.s,1],eps=eps.accuracy[,,all.s,1],along=4)
# 
# source('munge/bl.pre.process.R')
# source('src/bl.process.R')
```